---
title: "Homework Week 6"
author: "Yusang Hu 400177333"
date: "Oct 24 2019"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=15,fig.height=12)
```

## Introduction

Groceries is  a  data  set  contains  30  days  transaction  records  from  a  typical  local  grocery outlet. The data set is provided by Michael Hahsler, Kurt Hornik and Thomas Reutterer (2006).The data set contains 9835  transactions,  one  row  represents  one  transaction  and  columns  represent  the goods which are purchased in specific transaction,and 169 kinds of goods (variables). 

In a supermarket, deliberate product placements would not only saves customers' shopping time  but  also  helps to  increase  sales  for  some  products. Studying associations between each items is helpful to classify each items into suited aisle. In this assignment, I'll study which items can be classified into same aisle with 'whole milk'. Note that I'll only study top 20 items and top 20 association rules in this assignment since there are too many item types.

## Histogram

First, I'll draw a histogram to display the frequency of top 20 items. I choose histogram because it helps us to visualize the distribution of our data, and what I'm interested in is studying the most frequent item and the least frequent item.

The histogram tells us that:
1. 'whole milk' is the best-seller in 30 days, and 'domestic eggs' has the worst sales in top 20 items in 30 days.
2. Since I'm going to study which items can be classified into same aisle with 'whole milk', and 'whole milk' is the best-seller, therefore we can change the layout of supermarket by putting some bad sales items (i.e. domestic eggs, brown bread) with whole milk to increase sales volume of those bad sales items.

```{r}
library(ggplot2)
library(arules)
rm(list=ls())
#setwd("/Users/admin/Desktop/744")

## JD: This works, but it would be better to use relative paths. Since your script and .csv are in the same place, you can use Rstudio "set path to script location" and then just say
## groceries<-read.transactions("groceries.csv",sep=",")

groceries<-read.transactions("https://raw.githubusercontent.com/huyusang1031/Stat744/master/groceries.csv",sep=",")
itemFrequencyGGPlot <- function(x, topN) {
  library(tidyverse)
  x %>%
    itemFrequency(type="absolute") %>%
    sort %>%
    tail(topN) %>%
    as.data.frame %>%
    tibble::rownames_to_column() %>%
    ggplot(aes(reorder(rowname, `.`),`.`)) + 
    geom_col() + 
    coord_flip() + ## flip plot 90 degrees, this will make the ranking of frequency more clearer
    scale_y_continuous(position="right") + ## move y-axis to the top
    labs(x="Items") +
    ggtitle("Items Frequency Histogram For Top 20 Items") + ## add title
    theme(plot.title=element_text(face="bold",hjust=0.5)) ## change layout of title
}  
itemFrequencyGGPlot(groceries, 20)
```

## Parallel Coordinate Plot

The second plot I'll draw is a parallel coordinate plot. It is a plot based on the association rules between each items. I picked top 20 association rules where their support is larger than 0.006 and confidence is larger than 0.5.

I choose to use parallel coordinate plot because the correlations between items can be spotted easily. What I'm interested in is studying associations between items, for example, the topmost line shows us that when I have yogurt and frankfurter in my shopping cart, I'm highly likely to buy whole milk as well.

The parallel coordinate plot tell us that:
1. Since I'm going to study which items can be classified into the same aisle with 'whole milk', I'll concentrate on finding the item that have association with whole milk. The line starts with whole milk shows us that when I have whole milk and onions in my shopping cart, I'm highly likely to buy other vegetables as well.

JD: OK, but why are you more interested in the rule that starts with whole milk than the ones that have whole milk at rhs?

```{r}
library(arulesViz)
params<-list(support=0.006,confidence=0.5)
fit<-apriori(groceries,parameter=params) 
plot(fit[1:20],method="paracoord",control=list(reorder=TRUE))
```

## Conclusion

From the histogram and the parallel coordinate plot, I'll conclude that I'll put whole milk, vegetables and eggs in same aisle.

JD: Your other conclusion seemed to be onions. Why is it eggs now?

## Rubric

JD: Your histogram is fine, but the explanations don't make much sense to me. Interested in highest and lowest is not a difference between scales (any scale can be used for that). Histograms with bars should always be linear and anchored at zero anyway, because of the physical interpretation.

Histogram: I choose linear scale because differences between each items can easily be measured with the help of linear scale; I choose anchored at zero, because I'm interested in which item has highest frequency and lowest frequency, I'm not interested in small differences between each bar. I flipped the histogram over because x-axis labels are hard to read in default plot (i.e. items' name are vertically). My categorical variables are ordered from highest frequency to lowest frequency, so it's sensible.

JD: More explanation of the rules and the paracoord plot are required. I was really not able to figure out what the rules mean, why there are extra lines sticking out of the rhs, or how you reached your final conclusion.

Parallel Coordinate Plot: The x-axis scale is special in this plot, it is neither linear, log, logit or sqrt, instead, it represents the position of each association rule, where '2' and '1' means the left hand side of rules and 'rhs' means the right hand side of rules. I didn't reorder y-axis labels because I'm not interested in the ranking of items.

## Reference

Hahsler, M., Hornik, K., & Reutterer, T. (n.d.). Implications of Probabilistic Data Modeling for Mining Association Rules. From Data and Information Analysis to Knowledge Engineering Studies in Classification, Data Analysis, and Knowledge Organization, 598–605. doi: 10.1007/3-540-31314-1_73

Grade: 1.6/3.
