---
title: "Assignment Week 7"
author: "Yusang Hu 400177333"
output:
  html_document:
    df_print: paged
Date: Nov 5, 2019
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=10,fig.height=8)
```

## Data Set

mpg is a data set from package 'ggplot2'. In this assignment, I'll study the relations between miles per gallon on highway and engine displacement, and also type of car, number of cylinders.

## Why using scatter plot

I choose to use scatter plot because it makes the comparisons more straightforward, readers can easily understand many different kinds of information including the relationship between miles per gallon on highway and engine displacement, the concentrate distribution of each type of car, how is the type of a car relate to the number of cylinders. 

## Why using interactive plot

However, scatter plot can give informations about 'relation', but it could not give any informations about 'background' of a specific point. Therefore, I added interactive component to my plot. In addition to engine displacement, number of cylinders, miles per gallon on highway and type of a car, I added some informations about a car's 'background' including manufacturer of a car, model of a car, and year of production. The interactive plot helps readers better understand a specific car's (or point) properties, and also helps potential purchasers obtain all useful informations about their favorite car.

## How to make interactive components clear

To make interactive components clear, I:
1. Rename labels (i.e. displ, cyl, hwy, class) in tooltip
2. Add three more labels: manufacturer, model, year
3. show new name of 'displ','cyl','hwy' and 'class', and three new labels in tooltip

## Rubric

Scatter plot:
1. Color: the original order of colors is confusing and looks inaesthetic in visual, hence, I reordered the color of bubbles
2. Legend: the ordering of the legend is approximately match order on plot after I reordered the color of bubbles
3. Layout (i.e. title, labels): I added a title to my plot, and renamed x-axis and y-axis to make them more meaningful

Interactive Components:
1. tooltip: I renamed labels and added three more labels to the tooltip, and the ordering of informations in the tooltip is appropriate (i.e. first three are numerical information, last five are literal information).

```{r}
library(plotly)
library(ggplot2)
p1<-mpg%>%
  ggplot(aes(x=displ,y=hwy,text=paste('engine displacement:',displ,
                                      '<br>number of cylinders:',cyl,
                                      '<br>highway miles per gallon:',hwy,
                                      '<br>type of Car:',reorder(class,cyl)),label=manufacturer,
             label2=model,label3=year))+
  geom_point(aes(color=reorder(class,cyl),size=cyl))+
  labs(x="Engine displacement (in litres)",y="Highway miles per gallon",size="model")+
  scale_size(range=c(3,8))+
  ggtitle("Engine Displacement vs. Highway Efficiency")+
  theme(plot.title=element_text(face="bold",hjust=0.5),legend.title = element_blank())
ggplotly(p1,tooltip = c("text","label","label2","label3"))
```

**BMB: the size legend seems to have gotten lost in plotly. Since you have overlapping points, I might have considered setting `alpha`<1, and/or specifying a colour for the edge of the point (use `pch=21`).  Did you think about a colour scale and/or background colour different from the default?  It might be more efficient to reorder the class once. I'm not quite sure what average number of cylinders by class tells me ...?  Score: 2**

